---
title: "acpbvar: a complete workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{acpbvar: a complete workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment  = "#>",
    fig.width  = 7,
    fig.height = 5,
    dpi = 100
)
options(mcmc.pb = FALSE)
```

```{r}
library(acpbvar)
```

This vignette walks through the full estimation pipeline of `acpbvar`
on a small simulated VAR for which the data-generating process is known.
The objective is twofold: to illustrate the public interface end to end,
and to verify that the package recovers a known structural impact matrix
from a sign-restriction identification.

## 1. A structural VAR with a known sign pattern

The helper `simulate_var_example()` returns a simulated sample of three
variables (output, inflation, interest rate) generated by a structural
VAR(1) with three structural shocks (supply, demand, monetary policy).
The structural impact matrix is calibrated to deliver the canonical
macroeconomic sign pattern: a positive supply shock raises output and
lowers inflation; a positive demand shock raises both; a positive
monetary-policy shock raises the interest rate while lowering output and
inflation.

```{r}
sim <- simulate_var_example(T = 200, seed = 42)

Y <- sim$Y
head(Y)
```

The true structural impact matrix, against which posterior estimates will
be compared, is:

```{r}
impact_true <- sim$impact_true
dimnames(impact_true) <- list(sim$var_names, sim$shock_names)
round(impact_true, 3)
```

## 2. Equation-by-equation preparation

The A-form representation requires equation-by-equation response vectors
and regressor matrices. The regressors include the lagged variables, an
intercept, and (for equations beyond the first) the contemporaneous
variables ordered before the current equation, entered with a negative
sign so that the impact matrix is lower triangular by construction.

```{r}
p   <- 1L
dat <- build_yX(Y, p = p)

length(dat$y_list)
sapply(dat$X_list, dim)
```

Univariate AR fits supply the residual variances that enter the
asymmetric Minnesota prior as scaling factors.

```{r}
s2out <- build_s2(dat$y_list)
round(s2out$s2, 3)
```

## 3. Empirical Bayes hyperparameter selection

`optim_hyper()` selects the own-lag and cross-lag tightness parameters
\((\kappa_1, \kappa_2)\) by maximising the closed-form log marginal
likelihood under the asymmetric Minnesota prior. Nelder-Mead optimisation
runs in log-space to guarantee positivity.

```{r}
opt <- optim_hyper(
    y_list  = dat$y_list,
    X_list  = dat$X_list,
    p       = p,
    s2      = s2out$s2
)

opt$kappa1
opt$kappa2
opt$log_ml
opt$convergence
```

## 4. Posterior simulation

`post_sim()` wraps the full pipeline: it prepares the data, fits the AR
variances, runs Empirical Bayes for the tightness, builds the asymmetric
prior, and draws `R` independent posterior values of the structural
coefficients and variances, equation by equation (exact Normal-Inverse-Gamma
draws, no Markov chain).

```{r}
set.seed(20260514)
mcmc <- post_sim(
    Y     = Y,
    p     = p,
    R     = 1000,
    seed  = 20260514
)

names(mcmc)
mcmc$spec$kappa
mcmc$spec$log_ml
```

## 5. Structural identification through sign restrictions

The sign-restriction matrix encodes the three identifying patterns. Rows
index shocks (supply, demand, monetary); columns index variables (output,
inflation, interest rate). A value of `+1` requires a positive response
on impact, `-1` a negative response, and `0` leaves the response
unrestricted.

```{r}
S <- rbind(
    supply   = c( 1, -1,  0),
    demand   = c( 1,  1,  0),
    monetary = c(-1, -1,  1)
)
colnames(S) <- sim$var_names
S
```

`sign_restrict()` implements the Rubio-Ramirez, Waggoner and Zha (2010)
rotation algorithm: for each posterior draw of the reduced form,
candidate rotations are drawn from the Haar measure and screened
against all `2^n_id` sign flips; the first accepted rotation is stored.

```{r}
accepted <- sign_restrict(mcmc, S, n_id = 3L, max_tries = 1000L, verbose = FALSE)
length(accepted)
```

The diagnostic vector attached to the result records, for each accepted
draw, how many `Q` candidates were screened before acceptance.

```{r}
n_tries <- attr(accepted, "n_tries")
summary(n_tries)
```

### Partial identification

<!-- Da scrivere (Federica). Fatti: S con meno righe delle variabili = identificazione parziale;
     le colonne oltre n_id restano libere e senza nome; basta il default n_id = nrow(S). -->

```{r}
S_partial <- S[c("supply", "demand"), ]
S_partial

accepted_partial <- sign_restrict(mcmc, S_partial, max_tries = 1000L, verbose = FALSE)
length(accepted_partial)
```

<!-- Da scrivere. Fatti: una riga di S tutta zero viene rifiutata; motivo: shock dichiarato
     identificato senza restrizioni, il suo segno verrebbe dall'ordine dei flip, non da un'ipotesi. -->

### Restrictions beyond impact

<!-- Da scrivere (Federica). Fatti: finora i segni valgono solo all'impatto (h = 0); con `horizons = K`
     la stessa matrice S è imposta a tutti gli orizzonti 0..K (è il disegno di Uhlig 2005, segni per sei mesi);
     le risposte ai vari orizzonti sono calcolate per ogni rotazione candidata, quindi servono più tentativi. -->

```{r}
accepted_h3 <- sign_restrict(mcmc, S, horizons = 3L, max_tries = 5000L, verbose = FALSE)
length(accepted_h3)
summary(attr(accepted_h3, "n_tries"))
```

<!-- Da scrivere. Fatti: per segni diversi a orizzonti diversi S è un array n_id x n_var x (horizons + 1),
     una matrice per orizzonte; una fetta di zeri lascia libero quell'orizzonte; ogni shock deve avere
     almeno una restrizione da qualche parte. Esempio sotto: pattern pieno all'impatto, poi solo il tasso
     a h = 1 e solo l'output a h = 2. -->

```{r}
S_array <- array(0, dim = c(3, 3, 3),
                 dimnames = list(sim$shock_names, sim$var_names, paste0("h", 0:2)))
S_array[, , "h0"] <- S
S_array["monetary", "interest_rate", "h1"] <- 1
S_array["supply", "output", "h2"] <- 1
S_array

accepted_arr <- sign_restrict(mcmc, S_array, horizons = 2L, max_tries = 5000L, verbose = FALSE)
length(accepted_arr)
```

<!-- Da scrivere. Fatti: l'identificazione parziale funziona allo stesso modo, S o l'array hanno n_id righe
     e le colonne oltre n_id restano libere a ogni orizzonte. -->

### Two ways of accepting a draw

<!-- Da scrivere (Federica). Fatti: `method = "per_draw"` (default) cerca fino a max_tries rotazioni per ogni
     draw e tiene la prima buona: la posteriore della forma ridotta resta quella del modello, Q è uniforme
     dentro l'insieme identificato di ciascun draw. `method = "joint"` estrae una sola rotazione per draw e
     scarta il draw se non passa, come nel codice di replica di Chan (2022): i draw di forma ridotta vengono
     ripesati per la misura del loro insieme identificato; gli accettati sono molti meno, serve R più grande.
     Chan (2022, nota 13) descrive entrambe le implementazioni. -->

```{r}
accepted_joint <- sign_restrict(mcmc, S, method = "joint", verbose = FALSE)
length(accepted_joint)
attr(accepted_joint, "method")
```

### Reading the diagnostics

<!-- Da scrivere. Tre numeri da guardare: convergence, failed_draws, max(n_tries) vs max_tries. -->

```{r}
mcmc$spec$convergence
```

<!-- Da scrivere. Fatti: convergence è il codice di optim() per la ricerca di (kappa1, kappa2);
     0 = convergenza; altrimenti i kappa in spec sono dove si è fermato l'ottimizzatore e post_sim ha avvisato. -->

```{r}
attr(accepted, "failed_draws")
```

<!-- Da scrivere. Fatti: failed_draws = indici dei draw senza rotazione entro max_tries;
     scartarli è giusto se l'insieme identificato è vuoto; è un problema solo se il limite morde. -->

```{r}
max(attr(accepted, "n_tries"))
```

<!-- Da scrivere. Regola: max(n_tries) molto sotto max_tries => gli scartati sono a insieme vuoto;
     vicino a max_tries => il limite taglia draw buoni, alzare max_tries finché non succede più. -->

## 6. Impulse responses

`collect_irfs()` stacks the per-draw impulse responses into a single
four-dimensional array; `plot_irfs()` summarises the posterior into a
median path and a credible band.

```{r}
H    <- 12L
irfs <- collect_irfs(accepted, n = 3L, p = p, H = H)
dim(irfs)
```

```{r, fig.width = 7.5, fig.height = 6.5}
plot_irfs(
    irfs,
    var_names   = sim$var_names,
    shock_names = sim$shock_names,
    probs       = c(0.16, 0.84)
)
```

The impact responses (horizon zero) carry the sign restrictions by
construction: the supply column is positive on output and negative on
inflation; the demand column is positive on both; the monetary column is
negative on output and inflation, positive on the interest rate.

### Numbers rather than pictures

<!-- Da scrivere. Fatti: irfs è variabile x orizzonte x shock x draw; l'orizzonte h sta in posizione h+1;
     le sintesi a un orizzonte sono quantili sull'ultimo indice. -->

```{r}
h <- 4L
resp <- irfs[, h + 1, , ]                  # variable x shock x draw
med  <- apply(resp, c(1, 2), stats::median)
lo   <- apply(resp, c(1, 2), stats::quantile, probs = 0.16)
hi   <- apply(resp, c(1, 2), stats::quantile, probs = 0.84)
dimnames(med) <- dimnames(lo) <- dimnames(hi) <- list(sim$var_names, sim$shock_names)

round(med, 3)                               # posterior median at horizon 4
round(lo, 3)                                # 16th percentile
round(hi, 3)                                # 84th percentile
```

<!-- Da scrivere. Stessa costruzione su tutti gli orizzonti = percorso completo di una risposta. -->

```{r}
path <- apply(irfs[1, , 3, ], 1, stats::quantile, probs = c(0.16, 0.5, 0.84))
colnames(path) <- paste0("h", 0:H)
round(path, 3)                              # output after a monetary shock
```

## 7. Recovery of the true impact

Averaging the impact column across accepted draws gives a posterior
estimate of the structural impact matrix, which can be compared against
the true matrix used in the simulation.

```{r}
impact_post <- Reduce("+", lapply(accepted, `[[`, "impact")) / length(accepted)
dimnames(impact_post) <- list(sim$var_names, sim$shock_names)
round(impact_post, 3)
```

```{r}
round(impact_true, 3)
```

The posterior recovers the sign pattern of the true impact matrix on
every entry, and the magnitudes are in the same neighbourhood, with the
expected sampling and identification uncertainty. Sign restrictions
identify the impact matrix only up to an orthogonal transformation
within the admissible set, so exact numerical recovery is not the
relevant standard; the recovery of the qualitative pattern, and of the
order of magnitude of the entries, is.

## 8. Common mistakes

<!-- Da scrivere. Fatti: ogni funzione esportata controlla gli argomenti e si ferma con un messaggio
     che nomina l'argomento e dice cosa serviva; i chunk sotto hanno error = TRUE per mostrare i messaggi. -->

<!-- Da scrivere: data frame al posto di matrice. -->

```{r, error = TRUE}
post_sim(as.data.frame(Y), p = 1L, R = 100L)
```

<!-- Da scrivere: ordine dei lag non intero. -->

```{r, error = TRUE}
post_sim(Y, p = 1.5, R = 100L)
```

<!-- Da scrivere: S con numero di colonne sbagliato; S con un valore fuori da {-1, 0, 1}. -->

```{r, error = TRUE}
sign_restrict(mcmc, S[, 1:2])
```

```{r, error = TRUE}
S_bad <- S; S_bad[1, 1] <- 2
sign_restrict(mcmc, S_bad)
```

<!-- Da scrivere: n diverso da quello dei draw accettati. -->

```{r, error = TRUE}
collect_irfs(accepted, n = 2L, p = p, H = 12L)
```

## References

- Chan, J. C. C. (2022). Asymmetric conjugate priors for large Bayesian
  VARs. *Quantitative Economics*, 13(3), 1145-1169.
- Rubio-Ramirez, J. F., Waggoner, D. F., & Zha, T. (2010). Structural
  Vector Autoregressions: Theory of Identification and Algorithms for
  Inference. *Review of Economic Studies*, 77(2), 665-696.
