---
title: "acpbvar: a complete workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{acpbvar: a complete workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment  = "#>",
    fig.width  = 7,
    fig.height = 5,
    dpi = 100
)
options(mcmc.pb = FALSE)
```

```{r}
library(acpbvar)
```

This vignette walks through the full estimation pipeline of `acpbvar`
on a small simulated VAR for which the data-generating process is known.
The objective is twofold: to illustrate the public interface end to end,
and to verify that the package recovers a known structural impact matrix
from a sign-restriction identification.

## 1. A structural VAR with a known sign pattern

The helper `simulate_var_example()` returns a simulated sample of three
variables (output, inflation, interest rate) generated by a structural
VAR(1) with three structural shocks (supply, demand, monetary policy).
The structural impact matrix is calibrated to deliver the canonical
macroeconomic sign pattern: a positive supply shock raises output and
lowers inflation; a positive demand shock raises both; a positive
monetary-policy shock raises the interest rate while lowering output and
inflation.

```{r}
sim <- simulate_var_example(T = 200, seed = 42)

Y <- sim$Y
head(Y)
```

The true structural impact matrix, against which posterior estimates will
be compared, is:

```{r}
impact_true <- sim$impact_true
dimnames(impact_true) <- list(sim$var_names, sim$shock_names)
round(impact_true, 3)
```

## 2. Equation-by-equation preparation

The A-form representation requires equation-by-equation response vectors
and regressor matrices. The regressors include the lagged variables, an
intercept, and (for equations beyond the first) the contemporaneous
variables ordered before the current equation, entered with a negative
sign so that the impact matrix is lower triangular by construction.

```{r}
p   <- 1L
dat <- build_yX(Y, p = p)

length(dat$y_list)
sapply(dat$X_list, dim)
```

Univariate AR fits supply the residual variances that enter the
asymmetric Minnesota prior as scaling factors.

```{r}
s2out <- build_s2(dat$y_list)
round(s2out$s2, 3)
```

## 3. Empirical Bayes hyperparameter selection

`optim_hyper()` selects the own-lag and cross-lag tightness parameters
\((\kappa_1, \kappa_2)\) by maximising the closed-form log marginal
likelihood under the asymmetric Minnesota prior. Nelder-Mead optimisation
runs in log-space to guarantee positivity.

```{r}
opt <- optim_hyper(
    y_list  = dat$y_list,
    X_list  = dat$X_list,
    p       = p,
    s2      = s2out$s2
)

opt$kappa1
opt$kappa2
opt$log_ml
opt$convergence
```

## 4. Posterior simulation

`post_sim()` wraps the full pipeline: it prepares the data, fits the AR
variances, runs Empirical Bayes for the tightness, builds the asymmetric
prior, and draws `R` independent posterior values of the structural
coefficients and variances, equation by equation (exact Normal-Inverse-Gamma
draws, no Markov chain).

```{r}
set.seed(20260514)
mcmc <- post_sim(
    Y     = Y,
    p     = p,
    R     = 1000,
    seed  = 20260514
)

names(mcmc)
mcmc$spec$kappa
mcmc$spec$log_ml
```

## 5. Structural identification through sign restrictions

The sign-restriction matrix encodes the three identifying patterns. Rows
index shocks (supply, demand, monetary); columns index variables (output,
inflation, interest rate). A value of `+1` requires a positive response
on impact, `-1` a negative response, and `0` leaves the response
unrestricted.

```{r}
S <- rbind(
    supply   = c( 1, -1,  0),
    demand   = c( 1,  1,  0),
    monetary = c(-1, -1,  1)
)
colnames(S) <- sim$var_names
S
```

`sign_restrict()` implements the Rubio-Ramirez, Waggoner and Zha (2010)
rotation algorithm: for each posterior draw of the reduced form,
candidate rotations are drawn from the Haar measure and screened
against all `2^n_id` sign flips; the first accepted rotation is stored.

```{r}
accepted <- sign_restrict(mcmc, S, n_id = 3L, max_tries = 1000L, verbose = FALSE)
length(accepted)
```

The diagnostic vector attached to the result records, for each accepted
draw, how many `Q` candidates were screened before acceptance.

```{r}
n_tries <- attr(accepted, "n_tries")
summary(n_tries)
```

## 6. Impulse responses

`collect_irfs()` stacks the per-draw impulse responses into a single
four-dimensional array; `plot_irfs()` summarises the posterior into a
median path and a credible band.

```{r}
H    <- 12L
irfs <- collect_irfs(accepted, n = 3L, p = p, H = H)
dim(irfs)
```

```{r, fig.width = 7.5, fig.height = 6.5}
plot_irfs(
    irfs,
    var_names   = sim$var_names,
    shock_names = sim$shock_names,
    probs       = c(0.16, 0.84)
)
```

The impact responses (horizon zero) carry the sign restrictions by
construction: the supply column is positive on output and negative on
inflation; the demand column is positive on both; the monetary column is
negative on output and inflation, positive on the interest rate.

## 7. Recovery of the true impact

Averaging the impact column across accepted draws gives a posterior
estimate of the structural impact matrix, which can be compared against
the true matrix used in the simulation.

```{r}
impact_post <- Reduce("+", lapply(accepted, `[[`, "impact")) / length(accepted)
dimnames(impact_post) <- list(sim$var_names, sim$shock_names)
round(impact_post, 3)
```

```{r}
round(impact_true, 3)
```

The posterior recovers the sign pattern of the true impact matrix on
every entry, and the magnitudes are in the same neighbourhood, with the
expected sampling and identification uncertainty. Sign restrictions
identify the impact matrix only up to an orthogonal transformation
within the admissible set, so exact numerical recovery is not the
relevant standard; the recovery of the qualitative pattern, and of the
order of magnitude of the entries, is.

## References

- Chan, J. C. C. (2022). Asymmetric conjugate priors for large Bayesian
  VARs. *Quantitative Economics*, 13(3), 1145-1169.
- Rubio-Ramirez, J. F., Waggoner, D. F., & Zha, T. (2010). Structural
  Vector Autoregressions: Theory of Identification and Algorithms for
  Inference. *Review of Economic Studies*, 77(2), 665-696.
